\vfill World of Warcraft's New Way to Get You Hooked

### 2013-02-13 paper.tex
Deleted superfluous text
Others' Arguments section worked on.

% This is the ethical question you are interested in answering. It must be one simple sentence and lead to a yes/no answer. It needs to be very narrowly focused, specific, and not abstract at all. It must involve a detailed case in the general area of your interest. Following the research question, write a line or two about why your question is important or interesting to anyone, does the answer to anyone?

@@

\section{Research Question}
Is Blizzard's use of World of Warcraft: Starter Edition ethical? \\
World of Warcraft: Starter Edition could precipitate an ``outbreak'' of videogame addiction leading to terrible incidents. \cite{SnowlyDeath} Deaths and addiciton linked to World of Warcraft would cause a sharp decline in the sale of games. Blizzard would suffer and be forced to lay off software engineers working on its products.

@@

% This is where you gather and present the arguments made by others interested in the same or similar questions (or more general versions of your question specialized to your case) . Do not reveal your judgments, just repeat the arguments for in the best possible light from the arguer's perspective. Cover both sides of your question (the "yes" side and the "no" side) to get a complete picture of how others are thinking about it. Choose commentators with some reputation when possible, choose one or two "on the edge" when it adds to the discussion. Do not include any general ethical principles or SE Code in here unless explicitly utilized by the researchers in their arguments. Make each argument short and simple, do not use long quotes or explain more than is necessary to understand the arguments at a fairly abstract level. Cite all arguments to their sources.

% [10%] Othersʼ Arguments (500 Words Max)
% __ Summarize the main arguments others have made, citing sources for each one. Make each summary under a separate, descriptive subheading.
% __ Must cover the strongest arguments on both sides of the issue. Make them clear, concise and objective (no opinion as to their strength).
% __ Do not use bare ethical principles (including the Code) in this section - they are not "other's" arguments about your focus question.
% __ Summarize all arguments very briefly at the end of this section.

\section{Arguments}
\subsection{Arguments Affirmative}
\subsubsection{Differing Player Archetypes}
Many studies and surveys disregard the fact that different types of people play MMORPGs for different reasons and these varying archetypes lead to a majority of players who are not in danger of addiction. By not making mention of this fact, sweeping generalizations are made without basis. \cite{PlayerMotivations}
\subsubsection{More Time Online Does Not Mean Addiction}
Many people who are characterized as spending ``excessive'' time online do not show symptoms of addiction, rather they show signs of psychological fulfillment that would otherwise elude them due to disabilities, circumstances, or social discomfort. Some people who are isolated or autistic find comfort in the use of the internet and games to create social relationships they would not have otherwise. \cite{IsThereEvidenceOfInternetAddiction}
\subsubsection{Excessive Users Do Not Exhibit Addiction Symptoms}
Users that spend more time playing the game than average exhibit characteristics that show using MMORPGs for social aspects of the in-game world as opposed to symptoms of addiction. These users would just as likely find fun elsewhere. \cite{ExcessiveUseForSocialAspects}
\subsection{Arguments Negative}
\subsubsection{More Intense Play By A Younger Audience Leads to Addiction}
``MMORPGs are played more intensely than video games,'' and are played primarily by a younger audience. Persons under the age of 27 (i.e. the majority of players of MMORPGs) are far more susceptible to addiction due to a lack of self identity and self confidence. \cite{YoungerAudiencesGetAddicted}
\subsubsection{Increased Time Playing Leads to Life Problems}
The longer a person plays World of Warcraft, the likelier they are to experience issues in their lives. \cite{ProblemUsageAmongHighlyEngagedPlayers}
\subsubsection{Addictive Games Have a Social Responsibility}
Current efforts to warn players of online games' addictiveness are far too minimal and often non-existent. Warning labels should be printed on games and companies should contact users who play in excess of a set threshold. \cite{MMORPGSocialResponsibility}\\

Arguments in favor of my conclusion that World of Warcraft \emph{is} ethical often boil down to the fact that people are complex and a simple correlation of time spent playing to a person being psychologically addicted is too far of a stretch. People play for a variety of reasons, many of which are psychologically fulfilling. Arguments against my conclusion bring up that often these players have immature brains and a lack of social identity due to age, leading to increased risk of addiction. They also cite studies that show increased play times leading to life issues and that the companies has a social responsibility to at least warn consumers of possible psychological consequences.

@@

% (1) Do a simple analysis of how the SE Code would apply to your question. Does the "Preamble" of the Code say anything about the actors in your question? Would an objective person agree that the Code, by its own terms, claims to govern the ethics of your question? (2) Give a short abstract of the basics of your expected analysis for your paper. What is your projected conclusion and what are the ethical principles to support your conclusion? How do the principles support your conclusion? (You will not be held to this conclusion in the end. Changes are to be expected, so give your best, first shot now.)

% [70%] Your Analysis (About 3000 Words)
% __ Start with a paragraph showing why the SE Code applies to your focus question. Do not merely conclude, but show the reader by application of Code principles.
% __ Liberally utilize sub-headings organized by the relevant SE Code sections and particular arguments to delineate your lines of reasoning.
% __ All your analysis must be thoroughly supported by reason and logic.
% __ All claims must be supported by reputable primary sources and formal data.
% __ Have 2-4 distinct sections of the SE code utilized in your analysis.
%__ Code section 1 can be central to one piece of analysis only. For example, if you determine section 1.03 applies, then no other section 1 Code tenet may be used later as a section heading.
% __ If discussion about “public good” is used, there must be data to support it. Simply writing “it benefits the general public because it would make many people happy” is insufficient.
% __ Utilitarian, Rawlsian and Deontological analyses must be integrated into your analysis where appropriate. They should support your SE Code analysis, not be isolated from it. (Exceptions must be approved by Dr. Turner.)
% __ Class readings should be referenced as appropriate to support your Code analysis.
% __ There should be a clear cohesiveness to the analysis such that each argument logically flows into the next and the reader is logically convinced that your conclusion has strength. Your writing must explicitly provide substantial arguments through logic and data that will gain the respect of a skeptic.
% __ Do not define ethical systems in your analysis unless they are required and are very brief (a few sentences maximum).
% __ Opinions + Assumptions + Feelings > /dev/null

@@

Blizzard's World of Warcraft is a software system that runs on computers. \cite{WoWSystemReqs} Software systems are written by Software Engineers. \cite{ACMApprovesSECode} According to the Computer Society and the ACM, all Software Engineers should abide by the rules outlined in the the SE Code of Ethics. \cite{ACMApprovesSECode}\cite{SECode} Therefore, Blizzard's World of Warcraft Software Engineers are subject to “the Code.” \cite{SECode}

>>>>

@@

\section{Research Question}
Is Blizzard's use of World of Warcraft: Starter Edition ethical? \\
World of Warcraft: Starter Edition could precipitate an ``outbreak'' of videogame addiction leading to terrible incidents. \cite{SnowlyDeath} Deaths and addiction linked to World of Warcraft would cause a sharp decline in the sale of games. Blizzard would suffer and be forced to lay off software engineers working on its products.

@@

% [10%] Others' Arguments (500 Words Max)
% __ Summarize the main arguments others have made, citing sources for each one. Make each summary under a separate, descriptive subheading.
% __ Must cover the strongest arguments on both sides of the issue. Make them clear, concise and objective (no opinion as to their strength).
% __ Do not use bare ethical principles (including the Code) in this section - they are not "other's" arguments about your focus question.
% __ Summarize all arguments very briefly at the end of this section.

\section{Arguments}
\subsection{Arguments Affirmative}
\subsubsection{Differing Player Archetypes}
Many studies and surveys disregard the fact that different types of people play MMORPGs for different reasons and these varying archetypes lead to a majority of players who are not in danger of addiction. By not making mention of this fact, sweeping generalizations are made without basis. \cite{PlayerMotivations}
\subsubsection{More Time Online Does Not Mean Addiction}
Many people who are characterized as spending ``excessive'' time online do not show symptoms of addiction, rather they show signs of psychological fulfillment that would otherwise elude them due to disabilities, circumstances, or social discomfort. Some people who are isolated or autistic find comfort in the use of the internet and games to create social relationships they would not have cultivated otherwise. \cite{IsThereEvidenceOfInternetAddiction}
\subsubsection{Excessive Users Do Not Exhibit Addiction Symptoms}
Users that spend more time playing the game than average exhibit characteristics that show using MMORPGs for social aspects of the in-game world as opposed to symptoms of addiction. These users would just as likely find fun elsewhere. \cite{ExcessiveUseForSocialAspects}
\subsection{Arguments Negative}
\subsubsection{More Intense Play By A Younger Audience Leads to Addiction}
``MMORPGs are played more intensely than video games,'' and are played primarily by a younger audience. Persons under the age of 27 (i.e. the majority of players of MMORPGs) are far more susceptible to addiction due to a lack of self identity and self confidence. \cite{YoungerAudiencesGetAddicted}
\subsubsection{Increased Time Playing Leads to Life Problems}
The longer a person plays World of Warcraft, the likelier they are to experience issues in their lives. \cite{ProblemUsageAmongHighlyEngagedPlayers}
\subsubsection{Addictive Games Have a Social Responsibility}
Current efforts to warn players of online games' addictiveness are too minimal and often non-existent. Warning labels should be printed on games and companies should contact users who play in excess of a set threshold. \cite{MMORPGSocialResponsibility}\\

Arguments in favor of my conclusion that World of Warcraft \emph{is} ethical boil down to the fact that people are complex and a simple correlation of time spent playing to a person being psychologically addicted is too far of a stretch. People play for a variety of reasons, many of which are psychologically fulfilling. Arguments against my conclusion bring up that often these players have immature brains and a lack of social identity due to age, leading to increased risk of addiction. They also cite studies that show increased play times leading to life issues and that the companies has a social responsibility to at least warn consumers of possible psychological consequences.

@@

% [70%] Your Analysis (About 3000 Words)
% __ Start with a paragraph showing why the SE Code applies to your focus question. Do not merely conclude, but show the reader by application of Code principles.
% __ Liberally utilize sub-headings organized by the relevant SE Code sections and particular arguments to delineate your lines of reasoning.
% __ All your analysis must be thoroughly supported by reason and logic.
% __ All claims must be supported by reputable primary sources and formal data.
% __ Have 2-4 distinct sections of the SE code utilized in your analysis.
%__ Code section 1 can be central to one piece of analysis only. For example, if you determine section 1.03 applies, then no other section 1 Code tenet may be used later as a section heading.
% __ If discussion about "public good" is used, there must be data to support it. Simply writing "it benefits the general public because it would make many people happy" is insufficient.
% __ Utilitarian, Rawlsian and Deontological analyses must be integrated into your analysis where appropriate. They should support your SE Code analysis, not be isolated from it. (Exceptions must be approved by Dr. Turner.)
% __ Class readings should be referenced as appropriate to support your Code analysis.
% __ There should be a clear cohesiveness to the analysis such that each argument logically flows into the next and the reader is logically convinced that your conclusion has strength. Your writing must explicitly provide substantial arguments through logic and data that will gain the respect of a skeptic.
% __ Do not define ethical systems in your analysis unless they are required and are very brief (a few sentences maximum).
% __ Opinions + Assumptions + Feelings > /dev/null

@@

Blizzard's World of Warcraft is a software system that runs on computers. \cite{WoWSystemReqs} Software systems are written by Software Engineers. \cite{ACMApprovesSECode} According to the Computer Society and the ACM, all Software Engineers should abide by the rules outlined in the the SE Code of Ethics. \cite{ACMApprovesSECode}\cite{SECode} Therefore, Blizzard's World of Warcraft Software Engineers are subject to "the Code." \cite{SECode}
